Inferring Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-10-31 doc/attic/framework.tex
checked though up to testing on mocks

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Pangloss: Testing the reconstruction method
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Line of Sight Mass Reconstruction]
{Reconstructing the Lensing Mass in the Universe \\
from Photometric Catalog Data}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

\todo{PJM}{Make abstract match conclusions.}

Strong gravitational lenses can provide high precision cosmological distance
measurements; one of the most important sources of systematic error in these
estimates is the additional mass along the line of sight to the source. Weak
lensing perturbations by galaxies and groups in this beam can contribute an ``external
convergence'' $\kappa$ that dominates the  uncertainty in the inferred
distances; likewise, it may also contaminate estimates of the luminosity of
objects at high redshift (such as reionisation-era proto-galaxies, and type Ia
supernovae standard candles).  We characterise this uncertainty by marginalising
over a probability density function for $\kappa$: here, we investigate the use
of a simple halo model to estimate $\Pr(\kappa)$ given noisy estimates of the
photometric redshift and stellar mass of galaxies observed in optical imaging
surveys of the field in question. We use mock catalogs from the Millenium
Simulation and compare our reconstructed $\Pr(\kappa)$ to ray-traced $\kappa$ values,
we find that our model produces a precise estimator of $\kappa$ with only small biases.
For each lightcone the individual $\Pr(\kappa)$ from this reconstruction process are
typically 1.8 times more precise than the global distribution of $\kappa$ given perfect knowledge 
of the halo mass and redshift. After incorporating uncertainties on redshift and halo and stellar masses
we typically find that the reconstruction yields an improvement factor of 1.4. For each line of sight the
 $\Pr(\kappa)$ can be formed before the investment of light-curve monitoring: by following up only the 
most constrained lines of sight a factor of 2 improvement over the global $\pr(\kappax)$ should be 
achievable.

\end{abstract}

@@

\section{Introduction}
\label{sec:intro}

Every distant object we observe has had its apparent shape distorted,
and size and total brightness magnified (or demagnified) by a compound
weak gravitational lens constructed from all the mass distributed
between us and it. As \citet{Vale+White2003} and \citet{HilbertEtal2007}
showed, there are no empty lines of sight through our universe. This
fact makes gravitational lensing a potentially important source of
systematic error for any estimate of luminosity (or distance); this 
issue has been 
raised for \eg type Ia supernovae by
\citet[][]{Holz+Wald1998,Holz+Linder2005}, for gamma ray bursts by
\citet[][]{Oguri+Takahashi2006,Wang+Dai2011},  and for high redshift
galaxies by \citet{BradacEtal2009}, among others.

Along lines of sight containing strong gravitational lenses, the
perturbative effects of line of sight mass structure have been found to
be particularly important, with foreground and background structures
having a significant effect on the inferred lensing cross-section
\citep[\eg][]{WongEtal2012} and distance ratios
\citep[][]{DalalEtal2005}. Indeed, \citet{SuyuEtal2010} found that in
time delay lens cosmography the so-called ``external convergence''
$\kappa$ due to mass structures along the (unusually over-dense) line
of sight to the quadruply-imaged radio source B1608$+$656 was the
dominant source of uncertainty in their 6\% measurement of Hubble's
constant. The problem is that one simply doesn't know how magnified and
distorted the source has been, because the unlensed source is
unobservable. The external shear can potentially be recovered in the
mass-modelling of the lens, but the convergence is undetectable from the
image positions, shapes and relative fluxes; this is the famous {\emph{ 
mass sheet-degeneracy}} \citep[see e.g.][for details]{FalcoEtal1985}.

How can we correct for this perturbative lensing and make accurate
measurements through the Universe? When inferring global quantities
(such as the cosmological parameters), combining the results from a
large number of independent objects will tend to reduce the uncertainty
due to external convergence, as the ensemble-averaged magnification
approaches zero for an unbiased sample \citep[\eg][]{Holz+Linder2005}.
If the sample is biased, however, some residual systematic error will
remain even as the statistical uncertainty decreases. Lensing bias will
be present (to varying degrees) in any sample {\it selected} by
brightness (which includes magnification), or any other quantity that
depends on $\kappa$. This bias may be introduced either at the object
detection stage, or later on when making decisions about which objects
to study further.  Moreover, we expect these biases to be strongest in
the strong lens systems, where magnification is highest. Large samples
of lenses are expected to be discovered in the next decade in
ground-based optical imaging surveys \citep{Oguri+Marshall2010}: both
the detectability of these lenses,  and the selection of sub-samples of
them for further observation, could be sensitive to the external
convergence.

Several attempts have been made to do better than simply constructing a
large sample of objects, and averaging over the resulting convergence
distribution. The weak lensing effect can be detected observationally 
by measuring the
small distortions it induces on the images of many background galaxies
in the field \citep[see e.g.][for a review]{Schneider2006}. In a 30''
radius aperture around the strong lens Q0957$+$561,
\citet{NakajimaEtal2009} used deep HST imaging to measure a weak lensing
convergence of $0.17\pm0.06$, which was used in the time delay distance
measurement of \citep{FadelyEtal2009}. While this system has an
unusually high external convergence due to its location on the edge of a
cluster of galaxies, the precision with which the convergence at the
lens position can be estimated is indicative of what can be achieved
with high quality weak lensing shear data. We note that this uncertainty
is not dependent on the value of the convergence, but instead depends
primarily on the number density of weakly lensed, measurable galaxies
available.

In order to make more precise estimates of the convergence at a given
sky position, we need more information.   There have been some attempts
to reconstruct the local and line of sight mass distribution in strong
lens fields in some detail, based on photometric and spectroscopic
surveys of the galaxies nearby. These efforts have focused on  detecting
galaxy groups, ascertaining the visible galaxies' membership of those
groups, and assigning mass to the group and galaxy halos in order to
predict the external shear apparently required by the strong lens
models. Surveys by
\citet{Fassnacht+Lubin2002,AugerEtal2007,WilliamsEtal2006,MomchevaEtal2006}
all found groups of galaxies both hosting, and lying  along the line of
sight to, known gravitational lenses. Initial estimates of the effect of
these structures suggested the importance of assigning mass to both
galaxy and group halos, and to allowing for significant uncertainty in
the centroids of group halos.

The probability density function (PDF) for the external convergence
given our available data $\data$, $\Pr(\xkappa|\data)$,  is the  key
quantity we seek, since by marginalising over this distribution we can
correctly propagate the uncertainties due to line-of-sight perturbations
(as \citeauthor{FadelyEtal2009} did).  
%
\citet{GunnarssonEtal2006} investigated a related PDF,  for the
magnification, using a galaxy halo model. Applying empirical galaxy
scaling relations both to simulate mock catalogs and then to reconstruct
the line of sight mass distribution. While they did not explore highly
overdense lines of sight, or the impact of groups and clusters, they
found that under their simplifying assumptions, the dispersion in
apparent source brightness due to lensing magnification could be reduced
by a factor of two.
%
\citet{WongEtal2011} estimated the PDF for the external shear,
$\Pr(\gamma|\data)$ from their survey data, running similar Monte Carlo
reconstructions of the mass in nine lens fields based on
\citeauthor{WilliamsEtal2006} and \citeauthor{MomchevaEtal2006}'s
photometric and spectroscopic measurements of galaxies close to the line
of sight. They then  compared the resulting predicted shear
distributions with the external shear demanded by the strong lens model.
Their reconstructions showed most of the shear to have been generated by
bright galaxies within 2~arcminutes of the lens, and that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear. They found significant
discrepancies between the lens model shear and the Monte Carlo
predictions, but were unable to distinguish between the environment
modelling and the strong lens modelling as the cause of these
discrepancies.

These observational investigations have provided some very useful
insight into the problem of convergence estimation.  Coming from a more
theoretical direction, large numerical simulations have been used to
predict, from first principles, the distribution of likely external
shears and convergences along various lines of sight, including those
containing strong lens systems.  \citet{Holder+Schechter2003} and
\citet{DalalEtal2005} carried out ray-tracing calculations in N-body
simulations to estimate the distribution of external shear values. 
\citet{HilbertEtal2009} performed similar ray-tracing experiments
through the Millenium Simulation \citep{SpringelEtal2005}, generating a
predicted $\kappa$ at every position in a simulated sky.
\citet{HilbertEtal2009} found that \MS lines of sight with strong lenses
were not biased towards high $\kappa$, although selection functions for
discovering or using particular samples of strong lenses were not taken
into account. Addressing the possibility of using weak lensing shear to
reconstruct the magnification~$\mu$ that affects cosmography with
standard candles (and standard sirens), \citet{HilbertEtal2011} used the
\MS ray-tracing results to explore optimal mapping techniques and
quantify their performance. They found that deep weak lensing shear
surveys (with background galaxy number densities of 100 arcmin$^{-2}$)
would enable $\Pr(\mu|\data)$ to be estimated such that the distance
uncertainty per object would be reduced by  20\% on average, and by 40\%
in regions of low convergence, leading to the suggestion that weak
lensing maps be used to identify the (low convergence)  lines of sight
that give higher precision distance estimates.

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy overdensity, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed overdensity towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappa$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappa)$, which was then marginalised over when inferring the time
delay distance in this system. 
% B1608$+$656 is indeed a high $\kappa$
% line of sight, but it does not reside in a massive cluster environment.
In a companion paper to this one, Greene et al (submitted) investigate
improvements to this method by weighting the galaxy counts by galaxy
mass, and perpendicular distance to the line of sight, again using the
\MS mock catalogs and their associated $\kappa$ values to construct
$\Pr(\kappa|\data)$.

In this paper, we combine the halo model reconstruction approach of
\citeauthor{GunnarssonEtal2006} and \citeauthor{WongEtal2011}, with the
idea of calibrating to simulations from \citeauthor{SuyuEtal2010}, in
order to put the halo model for  estimating the weak lensing convergence
due to mass along the line of sight to a distant source into a sound
framework for evaluating this important source of systematic
uncertainty, as we look forward to ever increasing sample sizes and
statistical precision. Taking the $\kappa$ values from
\citeauthor{HilbertEtal2009}'s ray-tracing calculations as the ``truth''
that we have to recover, we use the \MS mock galaxy catalogs first to
{\it calibrate} the reconstruction, and account for unseen mass and
voids, and then to {\it test} the accuracy of the line of sight mass
reconstruction under this assumption. Probabilistically assigning mass
and redshift to every observed galaxy in a given observed field, we
generate Monte Carlo sample line of sight mass distributions, and so
construct the PDF $\Pr(\kappa|\data)$ for that field. We then emulate
the combination of many such PDFs to quantify the residual biases that
would be translated to the global (including cosmological) parameters.
In doing so, we aim to answer the following questions:

\item Faint galaxies, filaments and dark structures will not appear in
any photometric object catalog, but they will contribute convergence at
some level. How much of the total convergence comes from visible
galaxies, and how much effect do dark structures and voids have?

\item Can the true convergence be recovered from a calibrated halo model
reconstruction? What scatter and residual bias are induced by the
reconstruction process, and how might these be reduced in the future?

\item We have some choices to make when planning follow-up observations
of interesting sightlines. How should we select objects to achieve the
highest accuracy convergence reconstruction? Can such a selection be
made robustly, using the reconstruction results?

This paper is organized as follows. We review the relevant gravitational
lensing theory in~\Sref{sec:theory}, and the Millenium Simulation
ray-traced convergences and mock catalogs in \Sref{sec:MS}, before
introducing our simple reconstruction  model in \Sref{sec:model}. We
then test this model in two phases: first, in \Sref{sec:knownMh+z}, with
known redshift and halo mass for every galaxy in a lens field, in order
to quantify the irreducible uncertainty due to unseen mass, and second,
in \Sref{sec:obsMstar+z}, with realistic observational uncertainties on
the observed galaxies' stellar masses and redshifts. In
\Sref{sec:biases} we investigate the potential systematic error induced
by selecting a subset of sightlines. We discuss our results in
\Sref{sec:discuss} before concluding in \Sref{sec:conclude}.

Throughout this paper magnitudes are given in the AB system and
we adopt the Millenium Simulation's ``concordance'' parameters for our reference cosmology, \ie
$h=0.73$, $\Omega_m=0.25$ and $\Omega_\Lambda=0.75$, where the symbols indicate
the Hubble Constant in units of 100 km s$^{-1}$ Mpc$^{-1}$ and the matter and
dark energy density of the Universe in units of the critical density.
.%\citep[e.g.\ ][]{KomEtal09}.

\begin{figure*}
\includegraphics[width=\textwidth]{figs/viewofalightcone.eps}
\caption[magcut]{Four different views of a slightly over-dense \MS line
of sight, with a $\kappa$ of 0.03. 
{\it Top row, left:} The positions of galaxies projected on the sky. The area
of the circles is proportional to observed $i$-band flux; the brightest
object shown has an $i$ magnitude of 17.7. 
{\it Top row, centre:} The angular sizes of halos projected on the sky.
Red and blue regions lie within the NFW scale radius and virial radius
of each halo, respectively. There are essentially no empty lightcones. 
{\it Top row, right:} The individual $\kappa$ contributions of
each halo, assuming a \citet{BMO} profile and the \citet{Neto2007}
mass-concentration relation. Comparison of
this panel and the centre panel reveals the relative importance of
proximity to the line of sight.
{\it Bottom:} A view along the redshift axis, showing only halos with
$|x|<0.3$ arcmin. The area of the points is proportional to each halo's
mass: the most massive halo shown has $1.6\times10^{12}\Msun$. The
optical axis is shown by the dashed line, while the dotted lines mark
the lens and source planes for a B1608-like strong lens.}
\label{fig:lightcone}
\end{figure*}

@@

Convergence, or Ricci focusing, occurs when a gravitational lens focuses
the rays in a given bundle. This focusing can cause distant objects to
appear brighter, and larger than they would if the lens were removed.
The convergence from an isolated mass sheet at a position
$\bmath{\theta}$ on the sky is the ratio of the projected surface
mass density ($\Sigma(D_{\mathrm{od}}\bmath{\theta})$) \devided by the
critical surface mass density at the redshift of the mass sheet
($\Sigma_{\rm cr}$),
\be
\kappa(\bmath{\theta})= \frac {\Sigma(D_{\mathrm{od}}\bmath{\theta})}
                              {\Sigma_{\rm cr}(\zd,\zs)}
\ee
where 
\be 
\label{eq:sigcrit} 
\Sigma_{\rm cr}(\zd,\zs) \equiv \frac{c^2 D_{\rm os}}{4 \pi G D_{\rm od} D_{\rm ds}}.
\ee
and the $D$'s are the angular diameter distances between the objects
refered to in the subscripts: o is the observer, d is the deflector and
s is the source. If the surface mass density of the sheet exceeds
$\Sigma_{\rm cr}$, multiple images of the source can occur, otherwise
only one image will be observed, although that image will still be
perturbed relative to the unlensed case.

@@

Strong lenses sit in our real Universe -- they are not \truely isolated;
there is some mass along the line of sight coming from intervening
structures. These structures may be physically associated with the lens
e.g. other members of a group or cluster, or they may be the result of
chance alignments at any redshift.  Whilst it is rare for three galaxies
to line up well enough for both of the background sources to be strongly
lensed \citep{GavazziEtal2008,CollettEtal2012a}, the large size of dark
matter halos makes a partial alignment -- with the additional structure
acting as a perturbing weak lens --  a near certainty
\citep{Vale+White2003,HilbertEtal2007}. These additional structures
induce an external shear which can be recovered in the mass-modelling of
the lens, but they can also introduce a convergence that is undetectable
from the image positions, shapes and relative fluxes
\citep{FalcoEtal1985}.

If the intrinsic source brightness changes, the observed brightness in
each image does not change simultaneously: there will be a time delay,
$\Delta t$, between the brightness change in each image. Despite the
invariance of the image positions and fluxes, under the addition of an
external convergence (and appropriate re-scaling of the lens potential)
the relative Fermat potential between the images is {\it not} invariant,
and so the time delay changes. As such it is necessary to include
$\kappa$ in the lens modelling, if cosmological parameters are to be
estimated accurately and precisely from observed time delays
\citep{SuyuEtal2010}. Mass distributions that are physically associated with the lens galaxy
will affect the stellar dynamics of the lens galaxy, so that dynamical
observations can break the internal mass-sheet degeneracy by providing
an additional estimate of the lens galaxy's mass
\citep[e.g.,][]{SuyuEtal2010}. In this work our aim is to quantify the
lensing effect of massive objects, regardless of their position along
the line of sight.

If there is external convergence present that is not included in the
lens modeling, then the time delay distance -- inferred assuming $\kappa
= 0$ -- will be $(1-\kappa)$ less than the true value of $\Dt$:
\be 
\label{eq:MassSheet:Dtbias}
\Dt^{\rm{true}}=(1-\kappa) \Dt^{{\kappax = 0}}.
\ee
We can hence estimate the true distance if we have  with additional
knowledge of $\kappa$. Since $\kappa$ is typically small the absolute
uncertainty on the estimate of $\kappa$ \correspends to the fractional
uncertainty with which time-delay distances can be \infered.

The relationship between convergence $\kappa$ and magnification $\mu$ is
\bea 
\mu &=&       \frac{1}{(1-\kappa)^2 - |\gamma|^2}, \notag \\
    &\approx& (1+2\kappa)
\label{eq:MassSheet:mag}
\eea
where we have used the facts that the magnitude of the gravitational
shear at any point is typically comparable to the magnitude of the 
convergence, and that both are typically small. 
Under these assumptions, the fractional uncertainty in the
inferred luminosity of a magnified source will be approximately  twice
the absolute uncertainty in the convergence~$\kappa$. These
relationships will allow us to put our results on convergence estimation
in context.

@@

\section{The Millennium Simulation}
\label{sec:MS}

In order to test the accuracy of our convergence estimates, we need to
know the true convergence for each line of sight. We cannot use  real
lines of sight for this,  but must use simulated lines of sight instead.
In this section we briefly review the Millenium Simulation, the ray
tracing calculations that have been carried out in it, and the mock
galaxy catalogs that have been produced.

The \MS \citep{SpringelEtal2005} is a cosmological N-body simulation of
Dark Matter structures in a cubic region approximately 680~Mpc in
comoving size, followed from redshift 127 to the present day. With an
approximate halo mass resolution of $2\times10^{10}{\rm M}_{\odot}$
(corresponding approximately to a galaxy with luminosity $0.1L^{*}$), it
provides a detailed prediction for the distribution of dark structures
present in the Universe, under the assumptions of the $\Lambda$CDM model
of hierarchical structure formation, cosmological densities as given at
the end of \Sref{sec:intro}, and $\sigma_8 = 0.9$. 
Galaxy properties,
such as stellar masses, luminosities and colours, were assigned to the
simulated halos according to a semi-analytic model for galaxy formation
\citep{DeLucia+Blaizot2007}: the resulting predictions for the galaxy
luminosity function and correlation function match the observational
data very well.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Pangloss: Testing the reconstruction paper
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title[Line of Sight Mass Reconstruction]
{Inferring Lensing Mass in the Universe \\
from Photometric Catalog Data}

@@

@@

\begin{abstract} 
Strong gravitational lenses can provide high precision cosmological distance
measurements; one of the most important sources of systematic error in these
estimates is the additional mass along the line of sight to the source. Weak
lensing perturbations by galaxies and groups in this beam can contribute an ``external
convergence'' $\kappa$ that dominates the  uncertainty in the inferred
distances; likewise, it may also contaminate estimates of the luminosity of
objects at high redshift (such as reionisation-era proto-galaxies, and type Ia
supernovae standard candles).  We characterise this uncertainty by marginalising
over a probability density function for $\kappa$: here, we investigate the use
of a simple halo model to estimate $\Pr(\kappa)$ given noisy estimates of the
photometric redshift and stellar mass of galaxies observed in optical imaging
surveys of the field in question. We use mock catalogs from the Millenium
Simulation and compare our reconstructed $\Pr(\kappa)$ to ray-traced $\kappa$ values,
we find that our model produces a precise estimator of $\kappa$ with only small biases.
For each lightcone the individual $\Pr(\kappa)$ from this reconstruction process are
typically 1.8 times more precise than the global distribution of $\kappa$ given perfect knowledge 
of the halo mass and redshift. After incorporating uncertainties on redshift and halo and stellar masses
we typically find that the reconstruction yields an improvement factor of 1.4. For each line of sight the
 $\Pr(\kappa)$ can be formed before the investment of light-curve monitoring: by following up only the 
most costrained lines of sight a factor of 2 improvement over the global $\pr(\kappax)$ should be 
achievable.
\end{abstract}

@@

\section{Introduction}

Every distant object we observe has had its shape distorted, and size
and total brightness magnified (or demagnified) by a compound lens constructed
from all the mass distributed between us and it. As \citet{Vale+White2003} and \citet{HilbertEtal2007}
showed, there are no empty lightcones in our universe. This fact makes gravitational lensing a potentially important source of systematic error for any estimate of luminosity or distance, an issue
raised for \eg type Ia supernovae by \citet[][]{Holz+Wald1998,Linder+Holz2004}, for
gamma ray bursts by \citet[][]{Oguri+Takahashi2006,Wang+Dai2011}, 
and for high redshift galaxies by \citet{BradacEtal2009}, among others.

Along lines of sight containing strong gravitational lenses,
the effect has been found to be
particularly important, with foreground and background structures
having a significant effect on the inferred lensing cross-section
\citep[\eg][]{WongEtal2012} and distance ratios \citep[][]{DalalEtal2005}.
\citet{SuyuEtal2010} found that in time delay lens cosmography the
lensing effect of mass structures along the (unusually over-dense) line of
sight to the quadruply-imaged radio source B1608$+$656, the so-called ``external
convergence'' $\kappax$, was the dominant source of uncertainty in 
their 6\% measurement of Hubble's constant.

The problem is that one simply doesn't know
how magnified and distorted the source has been because the unlensed
source is unobservable. The external shear can potentially be recovered
in the mass-modelling of the lens, but the convergence is undetectable
from the image positions, shapes and relative fluxes; this is the
famous {\emph{  mass sheet-degeneracy}} \citep[see e.g.][for details]{FalcoEtal1985}.

How can we correct for this perturbative lensing and make
accurate measurements through the Universe? When inferring global quantities
(such as the cosmological parameters), averaging the results from a large
number of independent objects will tend to reduce the uncertainty, as the
ensemble-averaged convergence approaches zero for an unbiased sample
\citep[\eg][]{Linder+Holz2004}. If the sample is biased, however, some
residual systematic error will remain while the statistical uncertainty
decreases -- and this will be the case for any sample selected by the
brightness of the lensed object, or any other quantity that depends on
$\kappa$. Large
samples of lenses are expected to be discovered in the next decade in
ground-based optical imaging surveys \citep{Oguri+Marshall2010} -- but these
samples will be image separation-limited, due to the surveys' relatively poor
($\sim 1$~arcsec) image quality. There is a possibility that
they are somewhat biased towards high values of $\kappa$, although 
\citet{Fassnacht+Lubin2002} \todo{Anyone}{check this is the right Fassnacht paper} found no evidence of lenses lying along over-dense lines of sight.

To date, attempts at reconstructing the local and line of sight mass
distribution in strong lens fields (what \citeauthor{WongEtal2011} call ``the
full environment'') have focused on photometric and  spectroscopic surveys to
detect galaxy groups and ascertain galaxies' membership in order to understand
the external shear apparently required by the strong lens models. Surveys by
\citet{Fassnacht+Lubin2002,AugerEtal2007,WilliamsEtal2006,MomchevaEtal2006}
all found groups of galaxies both hosting and lying  along the line of sight
to known gravitational lenses. Initial estimates of the effect of these
structures suggested the importance of assigning mass to both galaxy and group
halos, and to allowing for significant uncertainty in the centroids of group
halos. The probability density function (PDF) for the external convergence for
that lens, $\Pr(\xkappa)$, is the target for any reconstruction analysis, and
is the key quantity for any ensemble analysis, since by marginalising over it
we can correctly propagate the uncertainties due to line-of-sight perturbations. Coming
closest to estimating this quantity from survey data, \citet{WongEtal2011} ran
Monte Carlo reconstructions of the mass in nine lens fields based on
\citeauthor{WilliamsEtal2006} and \citeauthor{MomchevaEtal2006}'s
photometric and spectroscopic measurements of galaxies close to the line of
sight, and compared the resulting predicted shear distributions with the
external shear demanded by the strong lens model. Their reconstructions show most
of the shear being generated by bright galaxies within 2~arcminutes of the
lens, and that both line of sight structures and the group of galaxies in each
lens plane contribute significant proportions of the shear. They find
significant discrepancies between the lens model shear and the Monte Carlo
predictions, and are unable to distinguish between the environment modelling
and the strong lens modelling as the cause of these discrepancies.

These observational investigations have provided very useful insight into the
problem.  Coming from a more theoretical direction, large numerical
simulations have been used to predict, from first principles, the distribution
of likely external shears and convergences along strong lens lines of sight. 
\citet{Holder+Schechter2003} and \citet{Dalal+Watson2004}\todo{phil}{Dalal+Watson2004 doesn't seem to exist} carried out
ray-tracing calculations in N-body simulations to estimate the  distribution
of external shear values.  \citet{HilbertEtal2009} performed similar ray-tracing
experiments through the Millenium Simulation \citep{SpringelEtal2005},
generating a predicted $\kappa$ at every position in a simulated sky. \citet{HilbertEtal2009}
found that \MS lines of sight with strong lenses were not biased towards high $\kappa$,
although selection functions for discovering a strong lens were not taken into account.
The first step towards using this excellent calibration resource was taken by
\citet{SuyuEtal2010}, who selected Millenium Simulation lines of sight that
 had comparable apparent galaxy
overdensity in a 45 arcsec radius aperture down to $i < 24.5$. This overdensity
could be compared to the overdensity of the B1608$+$656 line of sight
\citep{FassnachtEtal2011}, and hence an estimate of  $\Pr(\kappa)$ could be
made. B1608$+$656 is indeed a high $\kappa$ line of sight, but it does not reside
in a typical lens environment.

In this paper, we bring these observational and theoretical approaches
together and  use numerical simulations to {\it test and calibrate} a
reconstruction of the  external convergence. Ray-tracing through the
simulations gives a ``truth'' that we can attempt to recover.  In this paper
we use the Millenium Simulation mock galaxy catalogs and ray-traced lensing
effects to test the accuracy of line of sight mass reconstructions.
Probabilistically assigning mass and redshift to every observed galaxy in the
field around a time delay lens, we generate Monte Carlo sample line of sight
mass distributions, and estimate the external convergence due to each one.
By using the Millenium siumlation as a calibration tool, we can convert the
total contribution from halos, $\Pr(\kappah)$, into a PDF for the likely convergence
along the line of sight $\Pr(\kappa)$.

We aim to answer the following questions:
\begin{itemize}
\item Faint galaxies, filaments and dark structures will not appear in a photometric
object catalog, but they will contribute convergence at some level. How much of
the total external convergence in a time delay lens system comes from visible
galaxies? How does this change as a function of magnitude cut? 
\item Can the true (ray-traced) convergence be recovered from a calibrated halo model
reconstruction? What scatter and residual bias are induced by the reconstruction
process? Which aspects of the model dominate these uncertainties? 
\item When faced with a newly-detected lens, surrounded by galaxies on the sky,
we have some choices to make when planning follow-up observations.
Which are the most important galaxies, with regard to the external convergence
produced? Can nearby groups and clusters be straightforwardly accounted for? If
lenses with such massive structures nearby are discarded, what impact does that
have on the distance accuracy?
\end{itemize}

This paper is organized as follows. In~\Sref{sec:theory} we review the relevant
gravitational lensing theory, before introducing our simple reconstruction 
model in \Sref{sec:model}.  We then test this model in two phases: first, in
\Sref{sec:knownMh+z}, with known redshift and halo mass for every galaxy in a
lens field, in order to quantify the irreducible uncertainty due to unseen mass,
and second, in \Sref{sec:obsMstar+z}, with realistic observational uncertainties
on the field galaxies' stellar masses and redshifts. In \Sref{sec:biases} we
investigate potential systematic errors induced by selecting a subset of sightlines.
We discuss our results in \Sref{sec:discuss} before
concluding in \Sref{sec:conclude}.

Throughout this paper magnitudes are given in the AB system and
we adopt the Millenium Simulation's ``concordance'' parameters for our reference cosmology, \ie
$h=0.73$, $\Omega_m=0.25$ and $\Omega_\Lambda=0.75$, where the symbols indicate
the Hubble Constant in units of 100 km s$^{-1}$ Mpc$^{-1}$ and the matter and
dark energy density of the Universe in units of the critical density
.%\citep[e.g.\ ][]{KomEtal09}.

\begin{figure*}
\includegraphics[width=\textwidth]{figs/viewofalightcone.eps}
\caption[magcut]{4 different views of a lightcone around a typical line of sight. This line of sight is slightly overdense with a $\kappa$ of 0.03. Top left show the positions of galaxies, the area of the circles is proportional to observed $i$-band flux; the brightest object shown has an $i$ magnitude of 17.7. Top Centre show the angular sizes of halos on the sky; red regions are within the NFW scale radius of a halo and blue shows the virial radius of each halo - there are clearly no empty lightcones. Top right shows the individual $\kappa$ contributions of each halo, assuming the \citet{BMO} profile and the \citet{Neto2007} mass-concentration relation. Bottom a view along the redshift axis, showing only halos with $|x|<0.3$ arcmin; the area of the points is scaled by halo mass; the most massive halo shown has $1.6\times10^{12}\Msun$; the optical axis is the dashed line whilst the dotted lines are the lens and source planes for a B1608-like strong lens.}
\label{fig:lightcone}
\end{figure*}

@@

Convergence, or Ricci focusing, occurs when a gravitational lens focuses the rays in a raybundle.
This focussing can cause distant objects to appear brighter, and larger than they would if the lens
 were removed.
The convergence from an isolated mass sheet is the ratio of the projected surface mass density
($\Sigma(D_{\mathrm{od}}\bmath{\theta})$) devided by the critical surface mass density at the redshift of the mass sheet ($\Sigma_{\rm cr}$),
\be
\kappa(\bmath{\theta})= \frac {\Sigma(D_{\mathrm{od}}\bmath{\theta})}{\Sigma_{\rm cr}(z)}
\ee
where 
\be \label{eq:sigcrit} 
\Sigma_{\rm cr}(z) \equiv \frac{c^2 D_{\rm os}}{4 \pi G D_{\rm od} D_{\rm ds}}.
\ee
and the D's are the angular diameter distances between the objects refered to in the subscripts: o is the observer, d is the deflector and s is the source. If the surface mass density of the sheet exceeds $\Sigma_{\rm cr}$, multiple images of the source can
occur, otherwise only one image will be observed, although that image will still be perturbed
relative to the unlensed case.

@@

Strong lenses sit in our real Universe -- they are not truely isolated;
there is some mass along the line of sight coming from intervening structures.
These structures may be physically associated with the lens e.g. other members
of a group or cluster, or they may be the result of chance alignments at any redshift. 
Whilst it is rare for three galaxies to line up well enough for both of the background sources
to be strongly lensed \citep{GavazziEtal2008,CollettEtal2012a}, the shear scale of dark matter halos
makes a partial alignment -- with the additional structure acting as a perturbing weak lens -- 
a near certainty \citep{Vale+White2003,HilbertEtal2007}. These additional
structures induce an external shear which can be recovered in the mass-modelling
of the lens, but they can also introduce a convergence that
is undetectable from the image positions, shapes and relative fluxes; {\it the
mass sheet-degeneracy} \citep{FalcoEtal1985}.

If the intrinsic source brightness changes, the observed brightness in each image does not change
simultaneously - there is a time delay, $\Delta t$, between the brightness change in
each image. Despite the invariance of the image positions and fluxes, under the addition
of an external convergence (and appropriate re-scaling of the lens potential) the relative Fermat potential between the images is {\it not}
invariant, and so the time delay changes. As such it is necessary to include $\kappa$
in the lens modelling, if cosmological parameters are to be estimated accurately and
precisely from observed time delays \citep{SuyuEtal2010}. If there is any external convergence, that is not included in the lens
modeling, then the time delay distance -- inferred assuming $\kappa = 0$ -- will be
$(1-\kappa)$ less than the true value of $\Dt$.
\be 
\label{eq:MassSheet:Dtbias}
\Dt^{\rm{true}}=(1-\kappa) \Dt^{{\kappax = 0}}.
\ee
We can overcome this degeneracy with additional knowledge of $\kappa$.
Since $\kappa$ is typically small the absolute uncertainty on the estimate
of $\kappa$ correspends to the percentage uncertainty with which time-delay
distances can be infered.

For mass-sheets that are physically associated with the lens galaxy, the mass sheet
 affects the stellar dynamics of the
lens galaxy, so dynamical observations can break the internal
mass-sheet degeneracy by providing an additional estimate of the lens galaxy's mass
\citep[e.g.,][]{SuyuEtal2010}. In this work our aim is to quantify the lensing effect of mass-sheets that are not physically associated with the lens galaxy -- mass-sheets
lying infront or behind the lens -- and the uncertainties induced by reconstructing
those mass-sheets using simple prescriptions applied to potential astronomical observables.

@@

\section{Estimating external convergence: The halo model approximation}
\label{sec:model}
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

For real lines of sight, we would like to estimate the external convergence given all
data available. Typically this data is in the form of a photometric galaxy
catalog, containing positions, brightnesses, colours, and possibly spectra for each galaxy in the
field of the lens, down to some magnitude
limit. Since we can not observe the surface mass density of each halo
directy, we need some way of estimating
the mass of each halo in the catalog, so that we can compute its
contribution to the total $\kappax$ along the line of sight. This mass assignment recipe will be uncertain, but
can be calibrated with cosmological simulations.

Transforming the photometometric galaxy catalogue into a  catalogue of
halo masses, positions, and redshifts, will enable us to attempt a
reconstruction of the convergence induced by every halo near a line of sight.
This is the subject of Section \ref{subsec:halos}. We also need to account for
the mass in the Universe that is not associated with the galaxies we can see
-- this is the subject of Section \ref{subsec:voids}.
